---
title: "DATA607_Assignment 9"
author: "Dhanya Nair"
date: "2024-03-22"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Assignment:
The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

## Methodology
1. I signed up in NYT to create a account.
2. then i created a app and enabled the API " Most Popular API" as it provides the Popular articles on NYTimes.com.
3. I called this new app : NYT_API . This generated the API key fo my access.


## Loading Packages

```{r,warning=FALSE}
#to make HTTP requests like GET 
library(httr)


#jsonlite is JSON parser/generator optimized for the web ..used for fromJSON()
library(jsonlite)


library(dplyr)
```


## get data from NYT mostpopular/v2/shared for the most shared articles on NYTimes.com  for last 7 days for share type: facebook

```{r}
#get data from NYT mostpopular/v2/shared for the most shared articles on NYTimes.com  for last 7 days for share type: facebook
NYT_popular_Raw <- GET("https://api.nytimes.com/svc/mostpopular/v2/shared/7/facebook.json?api-key=eRN4OzxJAsN20SXKM5AR2pKBNfUVMN5Y") 

#Parse to text
NYT_popular_C <- content(NYT_popular_Raw, as = 'text') 


#in json format
NYT_popular_J <- fromJSON(NYT_popular_C)

# Convert JSON to DataFrame
NYT_popular_J   <- NYT_popular_J$results

#select the required columns
NYT_popular_shared <- subset(NYT_popular_J , select = c(url, published_date, type))

#display the dataframe
class(NYT_popular_shared)
head(NYT_popular_shared)


```

## get data from NYT most popular/v2/viewed for the most viewed articles on NYTimes.com  for last 7 days


```{r}


#get data from NYT most popular/v2/viewed for the most viewed articles on NYTimes.com  for last 7 days
NYT_mostviewed_raw <- fromJSON("https://api.nytimes.com/svc/mostpopular/v2/viewed/7.json?api-key=eRN4OzxJAsN20SXKM5AR2pKBNfUVMN5Y",flatten = TRUE)

#pick the results 
NYT_mostviewed <-NYT_mostviewed_raw$results 

#confirm its a data frame
class(NYT_mostviewed)

#display the data frame
head(NYT_mostviewed)


```

## Using HTTR2

```{r}
library(httr2)

#request the data through WebAPI 
req <- request("https://api.nytimes.com/svc/mostpopular/v2/viewed/7.json?api-key=eRN4OzxJAsN20SXKM5AR2pKBNfUVMN5Y")

#Send the request
resp <- req |> req_perform()



# resp_body_* function is used to access the body. fo JSON use resp_body_json
# when i wrote the code without simplifyVector = TRUE , it was in a list of lists and i was unable to convert to a dataframe. hence used simplifyVector = TRUE to simplify 

resp_json <- resp |> resp_body_json(simplifyVector = TRUE)




#pick just the results 
NYT_mostviewed_1 <-resp_json$results



#confirm its a data frame
class(NYT_mostviewed_1)

#display the data frame
head(NYT_mostviewed_1)

```

## Conclusion:
I am able to use httr and jsonlite packages to read data through WebAPI , persist and transform to a dataframe.
I am able to use httr2 packages to read data through WebAPI , persist and transform to a dataframe.

