---
title: "DATA607Assignment 5"
author: "Dhanya Nair"
date: "2024-02-23"
output: html_document
---

1. Create a CSV file, read the file and tidy the data
```{r}
library(tidyr)
library(dplyr)
library(knitr)
library(stringr)

#read the csv file without any header row
load <- read.csv('https://raw.githubusercontent.com/datanerddhanya/DATA607/main/Arrival_delays.csv',sep = ",", header= FALSE,stringsAsFactors=FALSE, quote = "\"")

# provide a name to the columns
load <-    rename(load,
    Airline  = V1 ,
    Arrivaltype = V2,
    Los_Angeles = V3,
    Phoenix = V4,
    San_Diego = V5 ,   
    San_Francisco = V6,
    Seattle = V7
  )
   

#delete the blank row and the row with text showing destination names
 load<- load[grep("^[a-z]" , load$Arrivaltype),]
  
#update the two rows with missing airlines
load$Airline[2]= "ALASKA" 
load$Airline[4]= "AMWEST" 

 
```

2. Transform the data
```{r}

#pivot the data.frame longer
load_final <- load |> 
  pivot_longer(
     cols = !(Airline:Arrivaltype) ,
     names_to = "Destinations", 
     values_to = "flight_count" ,
    values_drop_na = TRUE
   ) 

#remove comma from the flight count values
 load_final$flight_count<- as.numeric(gsub(",","",load_final$flight_count))
  
```

3. compare the arrival delays for the two airlines

AMWEST has more delayed flights than ALASKA Airlines.
However for Seattle destination, the arrival delays are more for Alaska Airlines than AMWEST.
```{r}


    load_final %>%
    group_by(Airline) %>%
    filter(Arrivaltype == "delayed")  %>%
   summarize(flight_count = sum(flight_count) )

load_final %>%
    group_by(Destinations,Airline) %>%
    filter(Arrivaltype == "delayed")  %>%
   summarize(flight_count = sum(flight_count), )

load_final %>%
    group_by(Airline,Arrivaltype) %>%
     summarize(flight_count = sum(flight_count) )

```

By plotting the data, i see that Alaska has recorded greater proportion of delayed flights than AMWEST

```{r}
library(ggplot2)


  ggplot( data = load_final , aes( x=Airline  , y=flight_count, fill = Arrivaltype )) + 
    geom_bar(stat="identity") 
 
```

