---
title: "DATA607Assignment3"
author: "Dhanya Nair"
date: "2024-02-11"
output: html_document
---
#1 provide code that identifies the majors that contain either "DATA" or "STATISTICS"
```{r}
library(dplyr)
college_majors <- read.csv('https://raw.githubusercontent.com/datanerddhanya/DATA607/main/COLLEGE_MAJORS.csv')
college_majors_data_statistics <- college_majors[grep("DATA|STATISTICS", college_majors$Major,ignore.case = TRUE),]
head(college_majors_data_statistics)

```

#2 Write code that transforms the data


```{r}
library(tidyverse)
fruit_list <- c("bell pepper", "bilberry", "blackberry", "blood orange",
          "blueberry", "cantaloupe", "chili pepper", "cloudberry",
          "elderberry", "lime", "lychee", "mulberry",
          "olive", "salal berry")
str_view(fruit_list)
fruit_newlist  <- paste0("c(",paste0(" ","\"",fruit_list,"\"", collapse = ",") , ")")
str_view(fruit_newlist)
```


#3 Describe, in words, what these expressions will match:

(.)\1\1 

"(.)(.)\\2\\1"

(..)\1

"(.).\\1.\\1"

"(.)(.)(.).*\\3\\2\\1" 
```{r}

str_view(words,"(.)\1\1")
str_view(words,"(.)(.)\\2\\1")
str_view(words,"(..)\1")
str_view(words,"(.).\\1.\\1")
str_view(words,"(.)(.)(.).*\\3\\2\\1")
```


1. This expression matches the single letter contained in the first parenthesis repeated thrice. 
2. This expression matches the single letter contained in the first and second parenthesis repeated by the single letter contained in the second and first parenthesis. 
3. This expression matches the two letters in the parenthesis repeated twice.
4. This expression matches the single letter contained in the first parenthesis repeating twice anywhere in the word with only one letter between them.
5. This expression matches the single letters contained in the first,second and third parenthesis repeated by the single letters contained in the third, second and first parenthesis  with zero or more letters in between them.


#4 Construct regular expressions to match words that:
Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
library(dplyr)
str_view(words, "^(.).*\\1$")
str_view(words, "(..).*\\1")
str_view(words, "(.).*\\1.*\\1")

```




